Concurrent Programming with CXL

### 2020-06-25 book/paper.tex
Drafted text for ABP protocol

A distributed system is a concurrent system in which a collection
of processes communicate by message passing, much the same as
in the actor model.
The most important difference between distributed and concurrent
systems is that the former takes \emph{failures} into account.
In this chapter, we will consider two actors, Alice and Bob.
Alice wants to send a sequence of application messages to Bob,
but the underlying network may lose messages.
The network does not re-order messages: when sending messages
$m_1$ and $m_2$ in that order, then if both messages are received,
$m_1$ is received before $m_2$.
Also, the network does not create messages out of nothing: if
message $m$ is received, then message $m$ was sent.

@@

We assume that there are two unreliable network channels: \texttt{s\_chan}
from Alice to Bob, and \texttt{r\_chan} from Bob to Alice.
Because of the network's properties, we can model each
channel as a variable that either contains a network message or nothing
(we use \texttt{()} in the model).

The \emph{Alternating Bit Protocol} is suitable for our purposes.
Alice and Bob each maintain a zero-initialized sequence number,
\texttt{s\_seq} and \texttt{r\_seq} resp.
Alice sends a network message to Bob containing an
application message as \emph{payload} and Alice's sequence number
as \emph{header}.
When Bob receives such a network message, Bob returns an
an \emph{acknowledgment} to Alice, which is a network message
containing the sequence number in the message that Bob received.

@@

>>>>

A distributed system is a concurrent system in which a collection
of processes communicate by message passing, much the same as
in the actor model.
The most important difference between distributed and concurrent
systems is that the former takes \emph{failures} into account,
including failures of processes and failures of shared memory.
In this chapter, we will consider two actors, Alice and Bob.
Alice wants to send a sequence of application messages to Bob,
but the underlying network may lose messages.
The network does not re-order messages: when sending messages
$m_1$ and $m_2$ in that order, then if both messages are received,
$m_1$ is received before $m_2$.
Also, the network does not create messages out of nothing: if
message $m$ is received, then message $m$ was sent.

@@

The \emph{Alternating Bit Protocol} is suitable for our purposes.
We assume that there are two unreliable network channels: one from Alice
to Bob and one from Bob to Alice.
Alice and Bob each maintain a zero-initialized sequence number,
\texttt{s\_seq} and \texttt{r\_seq} resp.
Alice sends a network message to Bob containing an
application message as \emph{payload} and Alice's sequence number
as \emph{header}.
When Bob receives such a network message, Bob returns an
an \emph{acknowledgment} to Alice, which is a network message
containing the sequence number in the message that Bob received.

@@

Because of the network's properties, we can model each
channel as a variable that either contains a network message or nothing
(we use \texttt{()} in the model).  Let \texttt{s\_chan} be the channel
from Alice to Bob and \texttt{r\_chan} the channel from Bob to Alice.
\texttt{net\_send(pchan, m, reliable)} models sending a message \texttt{m} to
\texttt{\^{}pchan}, where \texttt{pchan} is either \texttt{\&s\_chan}
or \texttt{\&r\_chan}.
The method places either \texttt{m} (to model a successful send)
or \texttt{()} (to model loss) in \texttt{\^{}pchan}.
The use of the \texttt{reliable} flag will be explained later.
\texttt{net\_recv(pchan)} models checking \texttt{\^{}pchan} for
the next message.

Method \texttt{app\_send(m)} tries \emph{retransmitting} \texttt{m} until
an acknowledgment is received.
Method \texttt{app\_recv(reliable)} returns the next successfully received
message.
Figure~\ref{fig:abptest} shows how the methods may be used to send and receive
a stream of \texttt{NMSGS} messages reliably.
It has to be bounded, because model checking requires a finite model.

Only the last invocation of \texttt{app\_recv(reliable)} is invoked with
\texttt{reliable == True}.  It causes the last acknowledgment to be sent
reliably.  It allows the receiver (Bob) to stop, as well as the sender (Alice)
once the last acknowledgment has been received.
Without something like this, either the sender may be left hanging waiting
for the last acknowledgment, or the receiver waiting for the last message.
